Supporting information: Copulas and their potential for ecology

### 2019-09-20 SupportingInformation.Rmd
edits done as suggested by DCR: see my response

<!--Now the copula figures that were not presented in the main text-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvClayton.pdf}
\caption[Example SC copulas]{Log-transformed pdfs (A-E) and samples (F-J) 
from example SC 
copulas. S is Spearman correlation; p is the value of the parameter for the SC 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively. The parameter range for the family is $p \in (0,\infty)$, upper-tail dependence is $2^{-1/p}$ and lower-tail
dependence is $0$.\label{figped_SC}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
\caption[Example G copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
Spearman correlation; $p$ is the value of the parameter for the G family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter range for the family is $p \in [1,\infty)$, lower-tail dependence is $0$ and upper-tail
dependence is $2-2^{1/p}$.\label{figped_G}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvGumbel.pdf}
\caption[Example SG copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
copulas. S is Spearman correlation; $p$ is the value of the parameter for the SG 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively. The parameter range for the family is $p \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/p}$.\label{figped_SG}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvJoe.pdf}
\caption[Example SJ copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
copulas. K is Spearman correlation; $p$ is the value of the parameter for the SJ 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively. The parameter range for the family is $p \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/p}$.\label{figped_SJ}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
\caption[Example F copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; $p$ is the value of the parameter for the F family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter range for the family is $p \in (0,\infty)$, lower-tail dependence is $0$ and upper-tail
dependence is $0$.\label{figped_F}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
\caption[Example SBB1 copulas]{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in (0,\infty)$
and $p_2 \in [1,\infty)$, lower-tail dependence is $2-2^{1/p_2}$ and upper-tail
dependence is $2^{-1/(p_1 p_2)}$.\label{figped_SBB1}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
\caption[Example SBB6 copulas]{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/(p_1 p_2)}$.\label{figped_SBB6}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
\caption[Example SBB7 copulas]{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in (0,\infty)$, upper-tail dependence is $2^{-1/p_2}$ and lower-tail
dependence is $2-2^{1/p_1}$.\label{figped_SBB7}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
\caption[Example SBB8 copulas]{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in (0,1]$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/p_1}$ of $p_2=1$ and $0$ otherwise.\label{figped_SBB8}}
\end{center}
\end{figure}

@@

N and F copula families 
are known to have symmetric tail dependence,
i.e., strengths of dependence in the lower and upper tails are the same, for all copulas
in these families [@nelsen2006_copula]. C copulas are known to have stronger lower- than 
upper-tail dependence, and SC copulas, being rotations of C copulas by 
$180^\circ$, have stronger upper- than lower tail dependence [@nelsen2006_copula].
Using the `iRho` function from the `copula` package, for each of the four families N, 
F, C, and SC, we obtained parameters for which the expected value of 
Spearman's $\rho$ was, separately, $0,0.1,0.2,...,0.9$. For each copula family and for each 
of the selected parameter values we generated  $n$ points, 100 times, and thereby 
computed 100 values of the 
statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$. This exercise 
was repeated, separately, 
for $n=1000$ and for $n=35$, these cases being chosen, respectively, to reflect 
asymptotic behavior of the 
statistics and their behavior on the small datasets which are common in some fields
of ecology. The mean and standard error of the three difference statistics were computed for 
each copula family, parameter value, and value of $n$, and results were plotted against 
Spearman's rho (Figs \ref{fig_stat_testing35} and \ref{fig_stat_testing1000}), 
the expectation being that the difference statistics would be positive for 
C copulas, negative for SC copulas (except for the case of $\rho=0$, a 
boundary case of these families corresponding to independence), and not significantly different 
from $0$ for F and N 
copulas.
We used a $t$-test with a significance threshold of $0.05$ to judge significance here.
These expectations were fulfilled (Figs \ref{fig_stat_testing35} and 
\ref{fig_stat_testing1000}).

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas. See Appendix  \ref{Test_npa_stat} for details. $n=35$. \label{fig_stat_testing35}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas. See Appendix  \ref{Test_npa_stat} for details. $n=1000$.\label{fig_stat_testing1000}}
\end{center}
\end{figure}

@@

<!--Display model selection results for soilCN data in a table-->
```{r tab_soilCNfit, echo=F, results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h, #longtable = T, 
             format = "latex",linesep = "",
      caption.short = "Model fitting results for soil C and N dataset",
      caption = "Results of fitting our 16 copula families to the soil C and N dataset. Copula name abbreviations are given in the main text. AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.\\label{tab_soilCNfit}",
      booktabs = TRUE)
```

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for soil C and N data. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$. Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis. \label{fig_soilCN_nonparam}}
\end{center}
\end{figure}

<!--Function to generate a table for npa results-summary with bivariate data, this one about asymmetry of tail dependence-->
```{r make_table_stat_fn, echo=F}

fractxt<-function(fracval,numsurrog){
  ptxt<-''
  if (fracval>.5*numsurrog){                        
    ptxt<-paste(">",fracval,sep='')
  } 
  if (fracval<=.5*numsurrog){
    ptxt<-paste("<",numsurrog-fracval,sep='')
  }
  return(ptxt)
}

@@

<!--Displaying a table with npa summary results for soilCN data-->
```{r tab_soilCN_asym,echo=F,results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
#tab_soilCN_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
             format = "latex",linesep = "",
      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons. A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
```

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_birds_fit, echo=F, results='markup'}
h<-BivMS_birds$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h,
             format='latex',linesep = "",
             caption.short = "Model fitting results for the bird body mass and BMR dataset",
             caption = "Results of fitting our 16 copula families to the bird body mass and BMR dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_bmr_birds_fit}",
             booktabs=T)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for bird body mass and BMR data. Same format as Fig. \ref{fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_birds_nonparam}}
\end{center}
\end{figure}

<!--Displaying a table with npa summary results for Birds' BMR data-->
```{r tab_bmr_birds_asym, echo=F, results="markup"}
tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
#tab_bmr_birds_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_bmr_birds_asym,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
             caption = "Bird body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_bmr_birds_asym}",
             booktabs=T)
```

<!--Display model selection results for BMR:mammals data in a table-->
```{r tab_bmr_mammals_fit, echo=F, results='markup'}
h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h,
             format='latex',linesep = "",
             caption.short = "Model fitting results for the mammal body mass and BMR dataset",
             caption = "Results of fitting our 16 copula families to the mammal body mass and BMR dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_bmr_mammals_fit}",
             booktabs=T)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for mammal body mass and BMR data. Same format as Fig. \ref{fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_mammals_nonparam}}
\end{center}
\end{figure}

<!--Displaying a table with npa summary results for Mammals' BMR data-->
```{r tab_bmr_mammals_asym, echo=F, results="markup"}
tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
#tab_bmr_mammals_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_bmr_mammals_asym,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
             caption = "Mammal body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure. \\label{tab_bmr_mammals_asym}",
             booktabs=T)
```

<!--Display model selection results for Cedar-creek data for year 2000 in a table-->
```{r tab_cc2000fit, echo=F, results='markup'}
h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h,
             format='latex',linesep = "",
             caption.short = "Model fitting results for the Cedar Creek dataset",
             caption = "Results of fitting our 16 copula families to the Cedar Creek dataset, year 2000. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_cc2000fit}",
             booktabs=T)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_4_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data, year 2000. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$. If the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.\label{fig_cc2000_nonparam}}
\end{center}
\end{figure}

>>>>

<!--Now the copula figures that were not presented in the main text-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvClayton.pdf}
\caption[Example survival Clayton copulas]{Log-transformed pdfs (A-E) and samples (F-J) 
from example survival Clayton 
copulas. S is Spearman correlation; p is the value of the parameter for the survival Clayton 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively. The parameter range for the family is $p \in (0,\infty)$, upper-tail dependence is $2^{-1/p}$ and lower-tail
dependence is $0$.\label{figped_SC}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
\caption[Example Gumbel copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Gumbel copulas. S is 
Spearman correlation; $p$ is the value of the parameter for the Gumbel family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter range for the family is $p \in [1,\infty)$, lower-tail dependence is $0$ and upper-tail
dependence is $2-2^{1/p}$.\label{figped_G}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvGumbel.pdf}
\caption[Example survival Gumbel copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example survival Gumbel 
copulas. S is Spearman correlation; $p$ is the value of the parameter for the survival Gumbel 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively. The parameter range for the family is $p \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/p}$.\label{figped_SG}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvJoe.pdf}
\caption[Example survival Joe  copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example survival Joe  
copulas. K is Spearman correlation; $p$ is the value of the parameter for the survival Joe  
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively. The parameter range for the family is $p \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/p}$.\label{figped_SJ}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
\caption[Example Frank copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Frank copulas. S is Kendall correlation; $p$ is the value of the parameter for the Frank family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter range for the family is $p \in (0,\infty)$, lower-tail dependence is $0$ and upper-tail
dependence is $0$.\label{figped_F}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
\caption[Example survival BB1 copulas]{Log-transformed pdfs for example survival BB1 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in (0,\infty)$
and $p_2 \in [1,\infty)$, lower-tail dependence is $2-2^{1/p_2}$ and upper-tail
dependence is $2^{-1/(p_1 p_2)}$.\label{figped_SBB1}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
\caption[Example survival BB6 copulas]{Log-transformed pdfs for example survival BB6 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/(p_1 p_2)}$.\label{figped_SBB6}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
\caption[Example survival BB7 copulas]{Log-transformed pdfs for example survival BB7 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in (0,\infty)$, upper-tail dependence is $2^{-1/p_2}$ and lower-tail
dependence is $2-2^{1/p_1}$.\label{figped_SBB7}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
\caption[Example survival BB8 copulas]{Log-transformed pdfs for example survival BB8 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in (0,1]$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/p_1}$ of $p_2=1$ and $0$ otherwise.\label{figped_SBB8}}
\end{center}
\end{figure}

@@

Normal and Frank copula families 
are known to have symmetric tail dependence,
i.e., strengths of dependence in the lower and upper tails are the same, for all copulas
in these families [@nelsen2006_copula]. C copulas are known to have stronger lower- than 
upper-tail dependence, and survival Clayton copulas, being rotations of C copulas by 
$180^\circ$, have stronger upper- than lower tail dependence [@nelsen2006_copula].
Using the `iRho` function from the `copula` package, for each of the four families Normal, 
Frank, Clayton, and survival Clayton, we obtained parameters for which the expected value of 
Spearman's $\rho$ was, separately, $0,0.1,0.2,...,0.9$. For each copula family and for each 
of the selected parameter values we generated  $n$ points, 100 times, and thereby 
computed 100 values of the 
statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$. This exercise 
was repeated, separately, 
for $n=1000$ and for $n=35$, these cases being chosen, respectively, to reflect 
asymptotic behavior of the 
statistics and their behavior on the small datasets which are common in some fields
of ecology. The mean and standard error of the three difference statistics were computed for 
each copula family, parameter value, and value of $n$, and results were plotted against 
Spearman's rho (Figs \ref{fig_stat_testing35} and \ref{fig_stat_testing1000}), 
the expectation being that the difference statistics would be positive for 
C copulas, negative for survival Clayton copulas (except for the case of $\rho=0$, a 
boundary case of these families corresponding to independence), and not significantly different 
from $0$ for Frank and Normal 
copulas.
We used a $t$-test with a significance threshold of $0.05$ to judge significance here.
These expectations were fulfilled (Figs \ref{fig_stat_testing35} and 
\ref{fig_stat_testing1000}).

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the Normal and Frank copulas. See Appendix  \ref{Test_npa_stat} for details. $n=35$. \label{fig_stat_testing35}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the Normal and Frank copulas. See Appendix  \ref{Test_npa_stat} for details. $n=1000$.\label{fig_stat_testing1000}}
\end{center}
\end{figure}

@@

@@

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_birds_fit, echo=F, results='markup'}
h<-BivMS_birds$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
colnames(h)<-c("copula","AIC","AICw","BIC","BICw","LT","UT")
h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC
knitr::kable(h_ord,
             format='latex',linesep = "",row.names = F, digits = c(4,2,2,4,4),
             caption.short = "Model fitting results for the bird body mass and BMR dataset",
             caption = "Results of fitting our 16 copula families to the bird body mass and BMR dataset. Analogous to Fig. \\ref{MT-tab_soilCNfit}, see caption of that figure for details. \\label{tab_bmr_birds_fit}",
             booktabs=T)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for bird body mass and BMR data. Same format as Fig. \ref{MT-fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_birds_nonparam}}
\end{center}
\end{figure}

<!--Display model selection results for BMR:mammals data in a table-->
```{r tab_bmr_mammals_fit, echo=F, results='markup'}
h<-BivMS_mammals$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
colnames(h)<-c("copula","AIC","AICw","BIC","BICw","LT","UT")
h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC
knitr::kable(h_ord,
             format='latex',linesep = "",row.names = F, digits = c(4,2,2,4,4),
             caption.short = "Model fitting results for the mammal body mass and BMR dataset",
             caption = "Results of fitting our 16 copula families to the mammal body mass and BMR dataset. Analogous to Fig. \\ref{MT-tab_soilCNfit}, see caption of that figure for details. \\label{tab_bmr_mammals_fit}",
             booktabs=T)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for mammal body mass and BMR data. Same format as Fig. \ref{MT-fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_mammals_nonparam}}
\end{center}
\end{figure}

<!--Display model selection results for Cedar-creek data for year 2000 in a table-->
```{r tab_cc2000fit, echo=F, results='markup'}
h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
colnames(h)<-c("copula","AIC","AICw","BIC","BICw","LT","UT")
h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC
knitr::kable(h_ord,
             format='latex',linesep = "",row.names = F, digits = c(4,2,2,4,4),
             caption.short = "Model fitting results for the Cedar Creek dataset",
             caption = "Results of fitting our 16 copula families to the Cedar Creek dataset, year 2000. Analogous to Fig. \\ref{MT-tab_soilCNfit}, see caption of that figure for details. \\label{tab_cc2000fit}",
             booktabs=T)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_4_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data, year 2000. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving Normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$. If the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.\label{fig_cc2000_nonparam}}
\end{center}
\end{figure}

<!--Function to generate a table for npa results-summary with bivariate data, this one about asymmetry of tail dependence-->
```{r make_table_stat_fn, echo=F}
fractxt<-function(fracval,numsurrog){
  ptxt<-''
  if (fracval>.5*numsurrog){                        
    ptxt<-paste(">",fracval,sep='')
  } 
  if (fracval<=.5*numsurrog){
    ptxt<-paste("<",numsurrog-fracval,sep='')
  }
  return(ptxt)
}

@@

<!--Displaying a table with npa summary results for soilCN data-->
```{r tab_biv_npa_summary,echo=F,results="asis"}
library(kableExtra)
library(dplyr)
library(tibble)

datatype <- tibble(Data = c("Soil CN","","","Bird masses and BMR","","","Mammal masses and BMR","","","Cedar Creek","",""))

tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
tab_biv_npa_summary<-rbind(tab_soilCN_asym,tab_bmr_birds_asym, tab_bmr_mammals_asym, tab_cc_asym_2000)

tab_biv_npa_summary<-cbind(datatype,tab_biv_npa_summary)
saveRDS(tab_biv_npa_summary,"./Results/stat_results/tab_biv_npa_summary.RDS")

tab_3biv_npa_summary<-tab_biv_npa_summary[c(4:12),]

knitr::kable(tab_3biv_npa_summary, booktabs = T, escape=F,
             format = "latex",linesep = "", row.names=F,
      caption.short = "Nonparametric asymmetry tests against an Normal-copula null model with birds, mammals and Cedar Creek data",
      caption = "Nonparametric asymmetry tests against an Normal-copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence). For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for $(0,0.25)$ and $(0.75,1)$ range of lower and upper bounds. Though the non-parametric result does not indicate significantly stronger lower tail dependence for Cedar Creek data, the 
greater than (>) sign before the numbers do show the opposite direction in tail-dependence compared to other bivariate datasets.\\label{tab_3biv_npa_summary}")
```
